Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity

\begin{proof}
\alain{preuve pas claire, à réécrire}
% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for $3$-counter machines.
%	This is due to the undecidability of termination for $2$-counter Minsky machines.
From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$
such that $M$ terminates on all initial states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.
%
Remark that $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$.
%
Based on this construction, the downward-closed set 
$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable (in $M'$)
% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates on all initial states; this last problem is undecidable~\cite{Min61}. 
Note that by chosing $s_0 \in Q' \times \mathop{\uparrow} (0,0,1)$,
deciding whether the downward-closed set 
$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable
from every state in $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1)$ 
is again equivalent to asking whether $M$ terminates on every initial state.
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

% \alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS in which
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state $ q({\bf 0})$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
  Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
\end{proof}

>>>>

\begin{proof}
\alain{j'ai commencé à  réécrire mais je n'ai pas fini car il y a des choses non expliquées que je ne suis pas sûr de comprendre...}
% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for $3$-counter machines.
\alain{pourquoi prendre 3 compteurs et pas 2 ? On ne comprend pas}
%	This is due to the undecidability of termination for $2$-counter Minsky machines.
From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$. We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.

We observe that $M$ terminates from all (initial) states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
\alain{pourquoi  ? On ne comprend pas}
%			with at least $1$ on its third counter. 
%
Remark that $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$.
\alain{pourquoi   $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ ?}
\alain{pourquoi  $Q' \times \mathop{\uparrow}(0,0,1)$ et pas $Q' \times \mathop{\uparrow}(0,0,0)$? On ne comprend pas}
%
\alain{l'ensemble $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ est redondant car il est égal à $Q' \times \{(m,n,0) \mid m,n \in \N\}=Q' \times \N^2 \times \{0\}$}
%
\alain{les ensembles $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ et $Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ sonts les mêmes et égaux à $Q' \times \N^2 \times \{0\}$. Pourquoi avoir deux notations différentes ?? ça complique inutilement la lecture. Et prendre la troisième, la plus simple $Q' \times \N^2 \times \{0\}$ qui n'apparait pas...}
%
%
Based on this construction, the downward-closed set 
$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable (in $M'$)
% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates from all (initial) states;
%
\alain{a déjà été dit qq lignes avant}
%
 this last problem is undecidable~\cite{Min61}. 
Note that by chosing $s_0 \in Q' \times \mathop{\uparrow} (0,0,1)$,
deciding whether the downward-closed set 
$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable
from every state in $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1)$ 
%
\alain{n'a-t-on pas $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1) = Q' \times \mathop{\uparrow} (0,0,1)$ ? a-t-on $Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1)) = Q' \times  (0,0,0)$ ? si oui pourquoi faire compliqué ?}
%
is again equivalent to asking 
%
\alain{asking: non}
%
whether $M$ terminates on every initial state.
%
\alain{je trouve ça pas clair}
%
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

% \alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS in which
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state $ q({\bf 0})$ if and only if
%
\alain{attention notations incohérentes pour (q,0,0,0) et $ q({\bf 0})$}
%
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
  Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
%
\alain{désolé, pas clair}
%
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
\end{proof}
